Molecular dynamics simulation of a driven synchronized particle

### 2021-03-30 manuscript/manuscript_ajp_version2.tex
updated some language in exercises

In the following
we explore the behavior of our model
with suggested problems for interested readers.
We discuss the molecular dynamics algorithm
and
numerical integration techniques in Problem~\ref{ex:euler}.
We explore changes in the particle motion
with parameter changes in Problem~\ref{ex:parameters}.
We include analytic solutions to  the 
linear drag equation in Problem~\ref{ex:reynolds}, 
the equation of motion in Problem~\ref{ex:n2l}.
%[xx see comment below xx]
We extend the model with
numerical problems 
to include finite temperature effects
in Problem~\ref{ex:brownian}.

\begin{problem}{Write your own MD code}
  \label{ex:euler}
  %[xx you didn't really discuss other integration algorithms, you just mentioned some names That's OK. Instead you might want to make this problem 1 and guide readers to write their own program. Then mention that a program is available, which you only mention in the abstract xx]

  A working example of this code is available
  in Ref.~\cite{supplemental}
  If you would like to write your own,
  the following guide
  will walk you through our choices.
  In the code excerpts below,
  most comments and optional details are removed for clarity.

@@

  \item[(a)] {\bf Initialization}
    For convenience,
    we defined a Python dictionary
    to contain the simulation constants
    that could easily be passed by reference
    to subroutines.
    We have left as an exercise to the reader the addition
    of the remaining parameters of the system.
    \begin{verbatim}
def set_parameters():           
    ```set simulation parameters...'''

@@

  \item[(b)]
    A key decision for any MD algorithm
    is how much information to save during and after
    the simulation.
    We define the following constants
    to manage the length of arrays
    containing data.

    In the subroutine {\bf single\_particle()}
    calculate the array lengths to hold
    position, velocity, and time data.
    Define a loop to 
    calculate the particle position and velocity through each
    molecular dynamics time step,
    calculated in the subroutine {\bf md\_step()}
    (which will be described in more detail in (c)).

@@

\item[(c)]

  To calculate the position of the particle we
  numerically integrate the equation of motion using
  the standard definition of velocity
  $\vec{v} = d\vec{r}/dt$ 
  using the 
  Euler method. 
  The equation of motion provides
  a direct calculation for the particle velocity,
  as demonstrated in Problem~\ref{ex:n2l}.  
  The Euler method is effective for solving linear ordinary differential  equations
  of the form
  $dy/dt = f(t,y(t))$ with the initial condition $y(t_0) = y_0$.
  The solution is calculated algorithmically
  by stepping in time through $n$ integer steps
  $t_n = t_0 + n \Delta t$.
  At each subsequent step the new
  value for $y$ is calculated as a  solution of a map using
  discrete times 
  $y_{n+1} = y_n + f(t_n,y_n)$.
  Apply the Euler method to %our equation of motion
  Eq.~\ref{eq:motion}
  %[DM - resolved xx refer to specific equation by label xx]
  to determine  the analytic expression
  for the position of a particle
  $y_n$ at the $nth$ time step.

  In our implementation,
  we created a subroutine to consider each
  type of force in the system (external drive, landscape, etc).
  If the particle moves beyond the limits
  $0 \le y < L$,
  it must be returned 
  using the definition of periodic boundary conditions
  in Sec.~\ref{sec:MD}.

@@

    #calculate the net force on the particle
    #(left as an exercise to the reader)

    #calculate the new position
    #(left as an exercise to the reader)

@@

  %The molecular dynamics
  %described above
  %is available in Ref.~\cite{supplemental}
  We use the Python programming
  language
  for educational purposes.
  To model additional particles,
  the authors recommend a more traditional 
  compiled coding language
  to perform the many operations associated
  with neighbor lists to compute particle interactions.

@@

To first order the viscosity is $\eta \sim 10^{-3}$\,Pa-s.\cite{Volpe2013}
Use reasonable values for the
  experimental analog of this system and show 
that the Reynolds number is small.
  In addition to the values listed in Table~\ref{tab:1}, 
  the liquid density 
  $\rho \sim 10^3$ kg/cm$^3$ is a reasonable
  first order approximation.\cite{asce}
  \end{problem}

@@

  \bibitem{supplemental}
    \url{<https://github.com/mcddanielle/AJP_project/>}.

>>>>

In the following
we explore the behavior of our model
with suggested problems for interested readers.
We discuss the molecular dynamics algorithm
and
numerical integration techniques in Problem~\ref{ex:euler}.
%DM slight rewrite to remove the redundant ``we'' tense
Changes to the parameters are described in 
%We explore changes in the particle motion
%with parameter changes in
Problem~\ref{ex:parameters}.
We include analytic solutions to  the 
linear drag equation in Problem~\ref{ex:reynolds}, 
the equation of motion in Problem~\ref{ex:n2l}.
%[DM working on it xx see comment below xx]
%DM rewrite of the following sentence - unclear as written/edited
We extend the numerical model %with
%numerical problems 
to include finite temperature effects
in Problem~\ref{ex:brownian}.

\begin{problem}{Write your own MD code}
  \label{ex:euler}

  %[xx you didn't really discuss other integration algorithms, you just mentioned some names That's OK. Instead you might want to make this problem 1 and guide readers to write their own program. Then mention that a program is available, which you only mention in the abstract xx]
  To calculate the position of the particle
  %DM add
  as a function of time
  %DM end
  we
  numerically integrate the equation of motion
  %DM add
  (Eq.~\ref{eq:motion})
  %DM end
  using
  the standard definition of velocity
  $\vec{v} = d\vec{r}/dt$ 
  using the 
  Euler method.
  %DM substituted direct reference for ``The equation of motion...''
  Eq.~\ref{eq:motion} provides
  a direct calculation for the particle velocity
  from the net force on the particle,
  as demonstrated in Problem~\ref{ex:n2l}.

    A working example of this code is available
  in Ref.~\cite{supplemental}
  If you would like to write your own,
  the following guide
  will walk you through our choices.
  In the code excerpts below,
  most comments and optional details are removed for clarity.
    %The molecular dynamics
  %described above
  %is available in Ref.~\cite{supplemental}
  We use the Python programming
  language
  for educational purposes.
  To model additional particles,
  the authors recommend a more traditional 
  compiled coding language
  to perform the many operations associated
  with neighbor lists to compute particle interactions.

@@

  \item[(a)] {\bf Initialization.}
    For convenience,
    we define a Python dictionary
    to contain the simulation constants
    that could easily be passed by reference
    to subroutines.
    We have left as an exercise to the reader the addition
    of the remaining parameters of the system.
    \begin{verbatim}
def set_parameters():           
    ```set simulation parameters...'''

@@

   \item[(b)] {\bf Time loop.}
     The Euler method is effective
     for solving linear ordinary differential  equations
     of the form
     $dy/dt = f(t,y(t))$ with the initial condition $y(t_0) = y_0$.
     The solution is calculated algorithmically
     by stepping in time through $n$ integer steps
     $t_n = t_0 + n \Delta t$.
     At each subsequent step the new
     value for $y$ is calculated as a  solution of a map using
     discrete times 
     $y_{n+1} = y_n + f(t_n,y_n)$.
     Apply the Euler method to %our equation of motion
     Eq.~\ref{eq:motion}
     %[DM - resolved xx refer to specific equation by label xx]
     to determine  the analytic expression
     for the position of a particle
     $y_n$ at the $nth$ time step.

     Add a $for$ loop
     to step through each integer timestep.
     With this loop, 
     control
     the flow of the program
     and retain information for the particle position and other
     properties
     as a function of time.
     Assume this information
     will be calculated a subroutine.

     We used the subroutine {\bf single\_particle()}
     calculate the array lengths to hold
     position, velocity, and time data.
     The lengths of these arrays will be affected by
     how much data you choice to save
     (see comment following sample code).
     Define a loop to 
     calculate the particle position and velocity through each
     molecular dynamics time step,
     calculated in the subroutine {\bf md\_step()}
     which will be described in more detail in (c).

@@

{\it Comment:}     
     A key decision for any MD algorithm
     is how much information to save during and after
     the simulation.
     %DM if going to emphasize this, have the reader do something with it
     %cut out - and make a comment rather than a statement going nowhere
     We define the following constants
     to manage the length of arrays
     containing data.
     We found in practice that for 
     short simulations times
     we could save all data.

\item[(c)] {\bf Position and force calculations.}
  %[DM moved sentences about general numerical solution to beginning
  %and those about time to (b)]
  In our implementation,
  we created a subroutine to consider each
  type of force in the system (external drive, landscape, etc).
  If the particle moves beyond the limits
  $0 \le y < L$,
  it must be returned 
  using the definition of periodic boundary conditions
  in Sec.~\ref{sec:MD}.

@@

    #calculate the net force on the particle
    vy = #(left as an exercise to the reader)

    #calculate the new position
    y += #(left as an exercise to the reader)

@@

@@

 %DM moved to end of paragraph
%To first order the viscosity is $\eta \sim 10^{-3}$\,Pa-s.\cite{Volpe2013}
Use reasonable values for the
  experimental analog of this system and show 
that the Reynolds number is small.
In addition to the values listed in Table~\ref{tab:1},
to first order
the viscosity is $\eta \sim 10^{-3}$\,Pa-s.\cite{Volpe2013}
and 
the liquid density 
$\rho \sim 10^3$ kg/cm$^3$.\cite{asce}
%DM removed
%is a reasonable
%first order approximation
  \end{problem}

@@

  \bibitem{supplemental}
    %not sure if the manuscript can remain in the repo - that probably violate AIP rules https://publishing.aip.org/resources/researchers/rights-and-permissions/author-licenses/
    %will make a DOI for the repo
    %https://guides.github.com/activities/citable-code/
    \url{<https://github.com/mcddanielle/AJP_project/>}.
